### 2016-12-20 plos_one_submission/Paper_v4.tex
PLoS submission

%\bibliographystyle{chicago}
\bibliographystyle{plos2009}

@@

% Begin ...
\begin{document}
\begin{titlepage}
{\par\centering\textbf{\Large {Reduced order modeling and analysis of the human complement system}}}
\vspace{0.05in}
{\par \centering \large{Adithya Sagar, Wei Dai$^{\#}$, Mason Minot$^{\#}$, Rachel LeCover, and Jeffrey D. Varner$^{*}$}}
\vspace{0.10in}
{\par \centering {Robert Frederick Smith School of Chemical and Biomolecular Engineering}}
{\par \centering {Cornell University, Ithaca NY 14853}}
\vspace{0.1in}
{\par \centering \textbf{Running Title:}~A reduced order model of complement}
\vspace{0.1in}
{\par \centering \textbf{To be submitted:}~\emph{Scientific~Reports}}
\vspace{0.5in}
{\par \centering $^{\#}$ Denotes equal contribution}
\vspace{0.1in}
{\par \centering $^{*}$Corresponding author:}
{\par \centering Jeffrey D. Varner,}
{\par \centering Professor, Robert Frederick Smith School of Chemical and Biomolecular Engineering,}
{\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853}
{\par \centering Email: jdv27@cornell.edu}
{\par \centering Phone: (607) 255 - 4258}
{\par \centering Fax: (607) 255 - 9166}
\end{titlepage}
\date{}
\thispagestyle{empty}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}
Complement is an important pathway in innate immunity, inflammation, and many disease processes.
However, despite its importance, there are few validated mathematical models of complement activation.
In this study, we developed an ensemble of experimentally validated reduced order complement models.
We combined ordinary differential equations with logical rules to produce a compact yet predictive model of complement activation.
The model,  which described the lectin and alternative pathways, was an order of magnitude smaller than comparable models in the literature.
We estimated an ensemble of model parameters from \textit{in~vitro} dynamic measurements of the C3a and C5a complement proteins.
Subsequently, we validated the model on unseen C3a and C5a measurements not used for model training.
Despite its small size, the model was surprisingly predictive.
Global sensitivity and robustness analysis suggested complement was robust to any single therapeutic intervention.
Only the simultaneous knockdown of both C3 and C5 consistently reduced C3a and C5a formation from all pathways.
Taken together, we developed a validated mathematical model of complement activation that was computationally inexpensive,
and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
The model described experimental data, and predicted the need for multiple points of therapeutic intervention to fully disrupt complement activation.

@@

Complement is an important pathway in innate immunity. It plays a significant role in inflammation, host defense as well as many disease processes.
Complement was discovered in the late 1880s where it was found to 'complement' the bactericidal activity of natural antibodies \cite{OG_COMPLEMENT_REF}.
However, research over the past decade has suggested the importance of complement extends beyond innate immunity.
For example, complement contributes to tissue homeostasis \cite{ricklin2010complement}.
It has also has been linked with several diseases including Alzheimers, Parkinson's, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{ricklin2007complement, rittirsch2008harmful}.
Complement also plays positive and negative roles in cancer; attacking tumor cells with altered surface proteins in some cases, while potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
Lastly, several other important biochemical systems are integrated with complement including the coagulation cascade, the autonomous nervous system and inflammation \cite{ricklin2013complement}.
Thus, complement is important in a variety of beneficial and potentially harmful functions in the body.
Despite its importance, there have been few approved complement specific therapeutics,
largely because of safety concerns and challenging pharmacokinetic constraints, however, progress is being made \cite{Ricklin:2013aa}.

@@

Pairwise parameter perturbations identified crosstalk within the complement model (Fig. \ref{fig-pairwise-analysis}).
We perturbed each pairwise combination of parameters by 10\%, and calculated the distance between the perturbed and nominal state for each parameter set in the ensemble.
We then clustered the mean response of each parameter combination based upon the euclidian distance between the perturbed and nominal states into low (green),
medium (red) and high (blue) response clusters. A low response (white) meant the parameter perturbations did not significantly change the system state compared with the nominal case.
Four of the 28 parameters (or approximately 14\% of the overall model parameters) were in the high response cluster (Fig. \ref{fig-pairwise-analysis}, blue cluster).
These parameters included the rate constant controlling the basal formation of C3b (\#12), C3a degradation (\#26) as well as the catalytic rate constant governing CP C3 convertase activity (\#22).
The only C5 related parameter in the high response group was the rate constant controlling the formation of CP C5 convertase (\#15).
Approximately, 36\%, or 10 of the 28 model parameters, were clustered in the medium impact cluster (Fig. \ref{fig-pairwise-analysis}, red cluster).
Three parameters (\#10,~\#1,~\#27) were especially important in this cluster;
The reaction order governing CP C3 convertase activity was important (\#10),
along with the rate constant controlling C4a and C4b formation from C4 in the lectin initiation pathway (\#1), and the constant controlling the inhibitory action of C4BP (\#27).
Lastly,  50\% of the model parameters were clustered in the low response cluster (Fig. \ref{fig-pairwise-analysis}, green cluster).
Many of these parameters influenced complement activation; for example, parameter \#23 (the CP C3 convertase saturation constant) was important, just not to the extent of other model parameters.
Pairwise synergistic interactions between parameters were also identified.
For example, in the high impact cluster, three synergistic relationships were identified, a single positive and two negative cases.
Parameters \#12 (rate constant governing basal C3b formation) and \#15 (formation of CP C5 conevertase) acted synergistically to increase the system response.
On the other hand, simultaneously changing parameters \#12 and \#22 or \#15 and \#26 decreased the system response relative to a single perturbation.
However, the most striking examples of synergy occurred in the medium impact cluster; for example, simultaneously increasing parameters \#13 (rate constant governing AP C3 convertase formation)
and \#19 (saturation constant governing AP C5 convertase activity) significantly changed the model state.
Changes in parameter \#3 (rate constant governing C2a and C2b formation from C2) showed both positive and negative synergistic effects
depending upon the other parameter that was perturbed.
Taken together, [FINISH ME].
However, sensitivity coefficients quantify how changes in parameters e.g., rate or saturation constants affect model performance.
To more closely simulate a clinical intervention e.g., administration of anti-complement inhibitors, we performed robustness analysis in the absence and presence of flow.

Robustness analysis in the absence of flow suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig. \ref{fig-robustness-analysis}).
Robustness coefficients quantify the response of a protein to a macroscopic structural or operational perturbation to a biochemical network.
Here, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5 with and without lectin initiator.
We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3, C5 or the combination of C3 and C5 by 50\%, 90\% and 99\%.
This would be conceptually analogous to the administration of a C3 inhibitor e.g., Compstatin alone or combination with Eculizumab (Fig. \ref{fig-SA}C).
The response of the complement model to different knock-down magnitudes was non-linear; a 90\% knock-down had an order of magnitude more impact than a 50\% knock-down.
As expected, a C5 knockdown had no effect on C3a formation for either the alternate (Fig. \ref{fig-robustness-analysis}A) or lectin pathways
(Fig. \ref{fig-robustness-analysis}B).
However, C3a and to a greater extent C5a abundance decreased with decreasing C3 concentration in the alternate pathway (Fig. \ref{fig-robustness-analysis}A).
This agreed with the sensitivity results; changes in AP C3-convertase formation affected the downstream dynamics of C5a formation.
Thus, if we only considered the alternate pathway, C3 alone could be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway.
Yet, when both pathways were activated, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}B);
even 1\% of the nominal C3 was able to generate enough AP/CP C5 convertase to maintain C5a formation.
Thus, the only reliable intervention that consistently reduced both C3a and C5a formation for all cases was a knockdown of both C3 and C5.
For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by an order of magnitude,
while C3a was reduced to a lesser extent (Fig. \ref{fig-robustness-analysis}B).

@@

Despite its importance, there has been a paucity of validated mathematical models of complement pathway activation.
To our knowledge, this study is one of the first complement models that combined multiple initiation pathways with experimental validation of important complement products like C5a.
However, there have been several theoretical models of components of the cascade in the literature.
Liu and co-workers modeled the formation of C3a through the classical pathway using 45 non-linear ODEs  \cite{liu2011computational}.
In contrast, in this study we modeled lectin mediated C3a formation using only five ODEs.
Though we did not model all the initiation interactions in detail, especially the cross-talk between the lectin and classical pathways,
we successfully captured C3a dynamics with respect to different concentrations of lectin initiators.
The model also captured the dynamics of C3a and C5a formed from the alternate pathway using only seven ODEs.
The reduced order model predictions of C5a were qualitatively similar to the theoretical complement model of Zewde et al., which involved over 100 ODEs \cite{zewde2016quantitative}.
However, we found that the C3a produced in the alternate pathway was nearly three orders of magnitude greater than the C5a generated.
While this was in agreement with the experimental data \cite{morad2015time},
it differed from the theoretical predictions made by Zewde et al., who showed C3a was eight orders of magnitude greater than the C5a concentration \cite{zewde2016quantitative}.
In our model, the time profile of both C3a and C5a generated changed with respect to the quantity of zymosan (the lectin pathway initiator).
In particular, the C3a peak time was directly proportional to initiator, while the lag phase for generation was inversely proportional to the initiator concentration.
Korotaevskiy et al. showed a similar trend using a theoretical model of complement, albeit for much shorter time scales \cite{korotaevskiy2009non}.
Thus, the reduced order complement model performed at least as well as existing larger mechanistic models, despite being significantly smaller.

Global analysis of the complement model suggested potentially important therapeutic targets.
Complement malfunctions are implicated in a spectrum of diseases, however the
development of complement specific therapeutics has been challenging \cite{ricklin2007complement,morgan2015complement}.
Previously, we have shown that mathematical modeling and analysis can be useful tools to estimate therapeutically important mechanisms \cite{Luan:2007aa,Nayak:2008aa,Tasseff:2010aa,Rice:2016aa}.
In this study, we analyzed a validated ensemble of reduced order complement models to better understand the strengths and weaknesses of the cascade.
In the presence of an initiator, C3a and C5a formation was sensitive to CP C3/C5 convertase assembly and activity, and to a lesser extent lectin initiation parameters.
Formation of the CP convertases can be inhibited by targeting upstream protease complexes like MASP-1,2 from the lectin pathway (or C1r, C1s from classical pathway).
For example, Omeros, a protease inhibitor that targets the MASP-2 complex, has been shown to inhibit the formation of downstream convertases \cite{schwaeble2011methods}.
Lampalizumab and Bikaciomab, which target factor B and factor D respectively, or naturally occurring proteins such as Cobra Venom Factor (CVF), an analogue of C3b, could also attenuate AP convertase formation \cite{vogel2004recombinant,katschke2012inhibiting, hu2013therapeutic}.
Removing supporting molecules could also destabilize the convertases.
For example, Novelmed Therapeutics developed the antibody, NM9401 against propedin, a small protein that stabilizes alternate C3 convertase \cite{bansal2014humanized}.
Lastly, convertase catalytic activity could be attenuated using small molecule protease inhibitors. All of these approaches are consistent with the results of the sensitivity analysis.
On the other hand, robustness analysis suggested C3a and C5a generation could only be significantly attenuated by modulating the free levels of C3 and C5.
The most commonly used anti-complement drug Eculizumab, targets the C5 protein \cite{morgan2015complement}.
Several other antibodies targeting C5 are also being developed; for example, LFG316 targets C5 in Age-Related Macular Degeneration \cite{roguska2014generation},
while Mubodina is used to treat Atypical Hemolytic-Uremic Syndrome (aHUS) \cite{melis2015complement}.
Other agents such as Coversin \cite{weston2014clinical} or the aptamer Zimura \cite{epstein2007complement} could also be used to knockdown C5.
The peptide inhibitor Compstatin and its derivatives are promising approaches for the inhibition of C3 \cite{Mastellos:2015aa}.
However, while the knockdown of C3 and C5 affect C3a and C5a levels downstream, the abundance, turnover rate and population variation of these proteins make them difficult targets \cite{sissons1977metabolism, swaak1982determination}.
For example, the eculizumab dosage must be significantly adjusted during the course of treatment for aHUS \cite{noris2014dynamics}.
A validated complement model, in combination with personalized pharmacokinetic models of immune system function, could be an important development for the field.

@@

\subsubsection*{Complement activation under flow conditions.}
We estimated the dynamics of complement activation under flow using a two-compartment model, with variable compartment volumes.
We considered a main compartment (m), and a wound compartment (w), where complement was activated in the wound compartment by the addition of a pathogenic surface.
In the main compartment, the balance for species $i$ ($x_{m,i}$) was given by:
\begin{equation}
	\frac{V_{m}}{\tau_{i}}\frac{dx_{m,i}}{dt} = \left(\sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x}_{m},\mathbf{\epsilon},\mathbf{k}\right)\right)V_{m}-k_{m,w,i}x_{m,i}+k_{w,m,i}x_{w,i}-Cx_{m,i}
\end{equation}where $V_{m}$ denotes the volume of the main compartment, $k_{p,q,i}$ denotes transfer constant governing the transfer of species $i$ from compartment $p$ to compartment $q$,
and $C$ denotes the clearance constant from the main compartment. We assumed complement factors were synthesized in the main compartment.
The balance governing species $i$ in the wound compartment was given by:
\begin{equation}
	\frac{V_{w}}{\tau_{i}}\frac{dx_{w,i}}{dt} = \left(\sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x}_{w},\mathbf{\epsilon},\mathbf{k}\right)\right)V_{w}-k_{w,m,i}x_{w,i}+k_{m,w,i}x_{m,i}-Bx_{w,i}
\end{equation}where $V_{w}$ denotes the volume of the wound compartment, and $B$ denotes the rate of blood loss from the wound compartment.
Lastly, because of the volume loss through clearance in the main compartment, and bleeding from the wound compartment, the volumes of each compartment were dynamic modeled:
\begin{eqnarray}
	\frac{dV_{m}}{dt} &=& I_{m}+F_{w,m} - F_{m,w} - C\\
	\frac{dV_{w}}{dt} &=& I_{w}+F_{m,w} - F_{w,m} - B
\end{eqnarray}where $I_{m},~I_{w}$ denote the rate of liquid input into the main and wound compartment.

@@

\subsection*{Complement model analysis.}
\subsubsection*{Global sensitivity analysis.}
We conducted global sensitivity analysis to estimate which parameters and species controlled the performance of the reduced order model.
We computed the total variance-based sensitivity index of each parameter relative to the training residual for the C3a/C5a alternate and C3a/C5a lectin objectives
using the Sobol method \cite{SOBOL_METHOD}.
The sampling bounds for each parameter were established from the minimum and maximum value for that parameter in the parameter ensemble.
We used the sampling method of Saltelli \textit{et al.} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
where $N$ was the number of trials per parameters, and $d$ was the number of parameters in the model \cite{saltelli2010variance}.
In our case, $N$ = 400 and $d$ = 28, so the total sensitivity indices were computed using 23,200 model evaluations.
The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \cite{SALIB}.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{./figs/Fig1_Schematic_v2.pdf}
\caption{Simplified schematic of the human complement system. The complement cascade is activated through three pathways:  the classical, the lectin, and the alternate pathways.
Complement initiation results in the formation of classical or alternative C3 convertases, which amplify the initial complement response and signal to the adaptive immune system by cleaving C3 into C3a and C3b. C3 convertases further react to form C5 convertases which catalyze the cleavage of the C5 complement protein to C5a and C5b.
C5b is critical to the formation of the membrane attack complex (MAC), while C5a recruits an adaptive immune response.}\label{fig-schematic}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{./figs/Figure2_Fits_final_v2.pdf}
\caption{Reduced order complement model training.
An ensemble of model parameters were estimated using multiobjective optimization from C3a and C5a measurements with and without zymosan \cite{morad2015time}.
The model was trained using C3a and C5a data generated from the alternative pathway (\textbf{A}--\textbf{B}) and lectin pathway initiated with 1 mg/ml zymosan (\textbf{C}--\textbf{D}).
The solid black lines show the simulated mean value of C3a or C5a for the ensemble, while the dark shaded region denotes the 99\% confidence interval of mean.
The light shaded region denotes the 99\% confidence interval of the simulated C3a and C5a concentration.
All initial conditions were assumed to be at their physiological serum levels unless otherwise noted.
}\label{fig-fit}
\end{figure}

>>>>

%\bibliographystyle{chicago}
\bibliographystyle{plos2015}

@@

% Begin ...
\begin{document}
\begin{titlepage}
{\par\centering\textbf{\Large {Reduced order modeling and analysis of the human complement system}}}
\vspace{0.05in}
{\par \centering \large{Adithya Sagar, Wei Dai$^{\#}$, Mason Minot$^{\#}$, Rachel LeCover, and Jeffrey D. Varner$^{*}$}}
\vspace{0.10in}
{\par \centering {Robert Frederick Smith School of Chemical and Biomolecular Engineering}}
{\par \centering {Cornell University, Ithaca NY 14853}}
\vspace{0.1in}
{\par \centering \textbf{Running Title:}~A reduced order model of complement}
\vspace{0.1in}
{\par \centering \textbf{To be submitted:}~\emph{PLoS~ONE}}
\vspace{0.5in}
{\par \centering $^{\#}$ Denotes equal contribution}
\vspace{0.1in}
{\par \centering $^{*}$Corresponding author:}
{\par \centering Jeffrey D. Varner,}
{\par \centering Professor, Robert Frederick Smith School of Chemical and Biomolecular Engineering,}
{\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853}
{\par \centering Email: jdv27@cornell.edu}
{\par \centering Phone: (607) 255 - 4258}
{\par \centering Fax: (607) 255 - 9166}
\end{titlepage}
\date{}
\thispagestyle{empty}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}
Complement is an important pathway in innate immunity, inflammation, and many disease processes.
However, despite its importance, there are few validated mathematical models of complement activation.
In this study, we developed an ensemble of experimentally validated reduced order complement models.
We combined ordinary differential equations with logical rules to produce a compact yet predictive model of complement activation.
The model,  which described the lectin and alternative pathways, was an order of magnitude smaller than comparable models in the literature.
We estimated an ensemble of model parameters from \textit{in~vitro} dynamic measurements of the C3a and C5a complement proteins.
Subsequently, we validated the model on unseen C3a and C5a measurements not used for model training.
Despite its small size, the model was surprisingly predictive.
Global sensitivity and robustness analysis suggested complement was robust to any single therapeutic intervention.
Only the simultaneous knockdown of both C3 and C5 consistently reduced C3a and C5a formation from all pathways.
Taken together, we developed a validated mathematical model of complement activation that was computationally inexpensive,
and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
The model described experimental data, and predicted the need for multiple points of therapeutic intervention to fully disrupt complement activation.

@@

Complement is an important pathway in innate immunity. It plays a significant role in inflammation, host defense as well as many disease processes.
Complement was discovered in the late 1880s where it was found to 'complement' the bactericidal activity of natural antibodies \cite{OG_COMPLEMENT_REF}.
However, research over the past decade has suggested the importance of complement extends beyond innate immunity.
For example, complement contributes to tissue homeostasis \cite{ricklin2010complement}.
It has also has been linked with several diseases including Alzheimers, Parkinson's, multiple sclerosis, schizophrenia, rheumatoid arthritis and sepsis \cite{Ricklin:2007aa, rittirsch2008harmful}.
Complement also plays positive and negative roles in cancer; attacking tumor cells with altered surface proteins in some cases, while potentially contributing to tumor growth in others \cite{sarma2011complement, ricklin2013complement}.
Lastly, several other important biochemical systems are integrated with complement including the coagulation cascade, the autonomous nervous system and inflammation \cite{ricklin2013complement}.
Thus, complement is important in a variety of beneficial and potentially harmful functions in the body.
Despite its importance, there have been few approved complement specific therapeutics,
largely because of safety concerns and challenging pharmacokinetic constraints, however, progress is being made \cite{Ricklin:2013aa}.

@@

Pairwise parameter perturbations identified crosstalk within the complement model (Fig. \ref{fig-pairwise-analysis}).
We perturbed each pairwise parameter combination by 10\%, and calculated the distance between the perturbed and nominal state for each parameter set in the ensemble.
We then clustered the mean response of each parameter combination based upon the euclidian distance between the perturbed and nominal states into low (green),
medium (red) and high (blue) response clusters. A low response (white) meant the parameter perturbations did not significantly change the system state compared with the nominal case.
Four of the 28 parameters (or approximately 14\% of the overall model parameters) were in the high response cluster (Fig. \ref{fig-pairwise-analysis}, blue cluster).
These parameters included the rate constant controlling the basal formation of C3b (\#12), C3a degradation (\#26) as well as the catalytic rate constant
governing CP C3 convertase activity (\#22).
The only C5 related parameter in the high response group was the rate constant controlling the formation of CP C5 convertase (\#15).
Approximately, 36\%, or 10 of the 28 model parameters, were clustered in the medium impact cluster (Fig. \ref{fig-pairwise-analysis}, red cluster).
Three parameters (\#10,~\#1,~\#27) were especially important in this cluster;
The reaction order governing CP C3 convertase activity was important (\#10),
along with the rate constant controlling C4a and C4b formation from C4 in the lectin initiation pathway (\#1), and the constant controlling the inhibitory action of C4BP (\#27).
Lastly,  50\% of the model parameters were clustered in the low response cluster (Fig. \ref{fig-pairwise-analysis}, green cluster).
Many of these parameters influenced complement activation; for example, parameter \#23 (the CP C3 convertase saturation constant) was important, just not to the extent of other model parameters.
Pairwise synergistic interactions between parameters were also identified.
For example, in the high impact cluster, three synergistic relationships were identified, a single positive and two negative cases.
Parameters \#12 (rate constant governing basal C3b formation) and \#15 (formation of CP C5 conevertase) acted synergistically to increase the system response.
On the other hand, simultaneously changing parameters \#12 and \#22 or \#15 and \#26 decreased the system response relative to a single perturbation.
However, the most striking examples of synergy occurred in the medium impact cluster; for example, simultaneously increasing parameters \#13 (rate constant governing AP C3 convertase formation)
and \#19 (saturation constant governing AP C5 convertase activity) significantly changed the model state.
Changes in parameter \#3 (rate constant governing C2a and C2b formation from C2) showed both positive and negative synergistic effects
depending upon the other parameter that was perturbed.
Taken together, sensitivity coefficients quantified how changes in parameters or parameter combinations affected model performance.
However, individual parameters e.g., rate or saturation constants are not easily druggable.
To more closely simulate a clinical intervention e.g., administration of anti-complement inhibitors, we performed knock-down analysis on the initial values of C3 and C5
in the absence and presence of flow.

Knock-down analysis in the absence of flow suggested there was no single intervention that inhibited complement activation in the presence of both initiation pathways (Fig. \ref{fig-robustness-analysis}).
Robustness coefficients quantify the response of a protein to a macroscopic structural or operational perturbation to a biochemical network.
Here, we computed how the C3a and C5a trajectories responded to a decrease in the initial abundance of C3 and/or C5 with and without lectin initiator.
We simulated the addition of different doses of anti-complement inhibitor cocktails by decreasing the initial concentration of C3, C5 or the combination of C3 and C5 by 50\%, 90\% and 99\%.
This would be conceptually analogous to the administration of a C3 inhibitor e.g., Compstatin alone or combination with Eculizumab (Fig. \ref{fig-SA}C).
The response of the complement model to different knock-down magnitudes was non-linear; a 90\% knock-down had an order of magnitude more impact than a 50\% knock-down.
As expected, a C5 knockdown had no effect on C3a formation for either the alternate (Fig. \ref{fig-robustness-analysis}A) or lectin pathways
(Fig. \ref{fig-robustness-analysis}B).
However, C3a and to a greater extent C5a abundance decreased with decreasing C3 concentration in the alternate pathway (Fig. \ref{fig-robustness-analysis}A).
This agreed with the sensitivity results; changes in AP C3-convertase formation affected the downstream dynamics of C5a formation.
Thus, if we only considered the alternate pathway, C3 alone could be a reasonable target, especially given that C5a formation was surprisingly robust to C5 levels in the alternate pathway.
Yet, when both pathways were activated, C5a levels were robust to the initial C3 concentration (Fig. \ref{fig-robustness-analysis}B);
even 1\% of the nominal C3 was able to generate enough AP/CP C5 convertase to maintain C5a formation.
Thus, the only reliable intervention that consistently reduced both C3a and C5a formation for all cases was a knockdown of both C3 and C5.
For example, a 90\% decrease of both C3 and C5 reduced the formation of C5a by an order of magnitude,
while C3a was reduced to a lesser extent (Fig. \ref{fig-robustness-analysis}B).
Taken together, these results suggested that both C3 and C5 inhibitors must be administered in the presence of both intiation pathways.

@@

There has been a paucity of validated mathematical models of complement pathway activation.
To our knowledge, this study is one of the first complement models that combined multiple initiation pathways with experimental validation of important complement products like C5a.
However, there have been several theoretical models of components of the cascade in the literature.
Liu and co-workers modeled the formation of C3a through the classical pathway using 45 non-linear ODEs  \cite{liu2011computational}.
In contrast, in this study we modeled lectin mediated C3a formation using only five ODEs.
Though we did not model all the initiation interactions in detail, especially the cross-talk between the lectin and classical pathways,
we successfully captured C3a dynamics with respect to different concentrations of lectin initiators.
The model also captured the dynamics of C3a and C5a formed from the alternate pathway using only seven ODEs.
The reduced order model predictions of C5a were qualitatively similar to the theoretical complement model of Zewde et al., which involved over 100 ODEs \cite{zewde2016quantitative}.
However, we found that the C3a produced in the alternate pathway was nearly three orders of magnitude greater than the C5a generated.
While this was in agreement with the experimental data \cite{morad2015time},
it differed from the theoretical predictions made by Zewde et al., who showed C3a was eight orders of magnitude greater than the C5a concentration \cite{zewde2016quantitative}.
In our model, the time profile of both C3a and C5a generated changed with respect to the quantity of zymosan (the lectin pathway initiator).
In particular, the C3a peak time was directly proportional to initiator, while the lag phase for generation was inversely proportional to the initiator concentration.
Korotaevskiy et al. showed a similar trend using a theoretical model of complement, albeit for much shorter time scales \cite{korotaevskiy2009non}.
Thus, the reduced order complement model performed at least as well as existing larger mechanistic models, despite being significantly smaller.

Global analysis of the complement model suggested potentially important therapeutic targets.
Complement malfunctions are implicated in a spectrum of diseases, however the
development of complement specific therapeutics has been challenging \cite{Ricklin:2007aa,morgan2015complement}.
Previously, we have shown that mathematical modeling and analysis can be useful tools to estimate therapeutically important mechanisms \cite{Luan:2007aa,Nayak:2008aa,Tasseff:2010aa,Rice:2016aa}.
In this study, we analyzed a validated ensemble of reduced order complement models to better understand the strengths and weaknesses of the cascade.
In the presence of an initiator, C3a and C5a formation was sensitive to CP C3/C5 convertase assembly and activity, and to a lesser extent lectin initiation parameters.
Formation of the CP convertases can be inhibited by targeting upstream protease complexes like MASP-1,2 from the lectin pathway (or C1r, C1s from classical pathway).
For example, Omeros, a protease inhibitor that targets the MASP-2 complex, has been shown to inhibit the formation of downstream convertases \cite{schwaeble2011methods}.
Lampalizumab and Bikaciomab, which target factor B and factor D respectively, or naturally occurring proteins such as Cobra Venom Factor (CVF), an analogue of C3b, could also attenuate AP convertase formation \cite{vogel2004recombinant,katschke2012inhibiting, hu2013therapeutic}.
Removing supporting molecules could also destabilize the convertases.
For example, Novelmed Therapeutics developed the antibody, NM9401 against propedin, a small protein that stabilizes alternate C3 convertase \cite{bansal2014humanized}.
Lastly, convertase catalytic activity could be attenuated using small molecule protease inhibitors. All of these approaches are consistent with the results of the sensitivity analysis.
On the other hand, robustness analysis suggested C3a and C5a generation could only be significantly attenuated by modulating the free levels of C3 and C5.
The most commonly used anti-complement drug Eculizumab, targets the C5 protein \cite{morgan2015complement}.
Several other antibodies targeting C5 are also being developed; for example, LFG316 targets C5 in Age-Related Macular Degeneration \cite{roguska2014generation},
while Mubodina is used to treat Atypical Hemolytic-Uremic Syndrome (aHUS) \cite{melis2015complement}.
Other agents such as Coversin \cite{weston2014clinical} or the aptamer Zimura \cite{epstein2007complement} could also be used to knockdown C5.
The peptide inhibitor Compstatin and its derivatives are promising approaches for the inhibition of C3 \cite{Mastellos:2015aa}.
However, while the knockdown of C3 and C5 affect C3a and C5a levels downstream, the abundance, turnover rate and population variation of these proteins make them difficult targets \cite{sissons1977metabolism, swaak1982determination}.
For example, the eculizumab dosage must be significantly adjusted during the course of treatment for aHUS \cite{noris2014dynamics}.
A validated complement model, in combination with personalized pharmacokinetic models of immune system function, could be an important development for the field.

@@

% \subsubsection*{Complement activation under flow conditions.}
% We estimated the dynamics of complement activation under flow using a two-compartment model, with variable compartment volumes.
% We considered a main compartment (m), and a wound compartment (w), where complement was activated in the wound compartment by the addition of a pathogenic surface.
% In the main compartment, the balance for species $i$ ($x_{m,i}$) was given by:
% \begin{equation}
% 	\frac{V_{m}}{\tau_{i}}\frac{dx_{m,i}}{dt} = \left(\sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x}_{m},\mathbf{\epsilon},\mathbf{k}\right)\right)V_{m}-k_{m,w,i}x_{m,i}+k_{w,m,i}x_{w,i}-Cx_{m,i}
% \end{equation}where $V_{m}$ denotes the volume of the main compartment, $k_{p,q,i}$ denotes transfer constant governing the transfer of species $i$ from compartment $p$ to compartment $q$,
% and $C$ denotes the clearance constant from the main compartment. We assumed complement factors were synthesized in the main compartment.
% The balance governing species $i$ in the wound compartment was given by:
% \begin{equation}
% 	\frac{V_{w}}{\tau_{i}}\frac{dx_{w,i}}{dt} = \left(\sum_{j = 1}^{\mathcal{R}}\sigma_{ij}r_{j}\left(\mathbf{x}_{w},\mathbf{\epsilon},\mathbf{k}\right)\right)V_{w}-k_{w,m,i}x_{w,i}+k_{m,w,i}x_{m,i}-Bx_{w,i}
% \end{equation}where $V_{w}$ denotes the volume of the wound compartment, and $B$ denotes the rate of blood loss from the wound compartment.
% Lastly, because of the volume loss through clearance in the main compartment, and bleeding from the wound compartment, the volumes of each compartment were dynamic modeled:
% \begin{eqnarray}
% 	\frac{dV_{m}}{dt} &=& I_{m}+F_{w,m} - F_{m,w} - C\\
% 	\frac{dV_{w}}{dt} &=& I_{w}+F_{m,w} - F_{w,m} - B
% \end{eqnarray}where $I_{m},~I_{w}$ denote the rate of liquid input into the main and wound compartment.

@@

\subsubsection*{Global sensitivity analysis.}
We conducted global sensitivity analysis to estimate which parameters and species controlled the performance of the reduced order model.
We computed the total variance-based sensitivity index of each parameter relative to the training residual for the C3a/C5a alternate and C3a/C5a lectin objectives
using the Sobol method \cite{SOBOL_METHOD}.
The sampling bounds for each parameter were established from the minimum and maximum value for that parameter in the parameter ensemble.
We used the sampling method of Saltelli \textit{et al.} to compute a family of $N\left(2d+2\right)$ parameter sets which obeyed our parameter ranges,
where $N$ was the number of trials per parameters, and $d$ was the number of parameters in the model \cite{saltelli2010variance}.
In our case, $N$ = 400 and $d$ = 28, so the total sensitivity indices were computed using 23,200 model evaluations.
The variance-based sensitivity analysis was conducted using the SALib module encoded in the Python programming language \cite{SALIB}.

@@

\begin{figure}[ht]
\centering
%\includegraphics[width=1.0\textwidth]{./figs/Fig-1_Schematic_v2.pdf}
\caption{Simplified schematic of the human complement system. The complement cascade is activated through three pathways:  the classical, the lectin, and the alternate pathways.
Complement initiation results in the formation of classical or alternative C3 convertases, which amplify the initial complement response and signal to the adaptive immune system by cleaving C3 into C3a and C3b. C3 convertases further react to form C5 convertases which catalyze the cleavage of the C5 complement protein to C5a and C5b.
C5b is critical to the formation of the membrane attack complex (MAC), while C5a recruits an adaptive immune response.}\label{fig-schematic}
\end{figure}

\begin{figure}[ht]
\centering
%\includegraphics[width=1.0\textwidth]{./figs/Fig-2_Fits_final_v2.pdf}
\caption{Reduced order complement model training.
An ensemble of model parameters were estimated using multiobjective optimization from C3a and C5a measurements with and without zymosan \cite{morad2015time}.
The model was trained using C3a and C5a data generated from the alternative pathway (\textbf{A}--\textbf{B}) and lectin pathway initiated with 1 mg/ml zymosan (\textbf{C}--\textbf{D}).
The solid black lines show the simulated mean value of C3a or C5a for the ensemble, while the dark shaded region denotes the 99\% confidence interval of mean.
The light shaded region denotes the 99\% confidence interval of the simulated C3a and C5a concentration.
All initial conditions were assumed to be at their physiological serum levels unless otherwise noted.
}\label{fig-fit}
\end{figure}
